---
title: "main"
output: html_document
---

### Read the data
```{r}
library(readr)
library(dplyr)
library(ggplot2)

data.set <- read.table("dataLA.prn", header=TRUE, na.strings="*")
data.set
```

### KeyStroke Distributions
```{r}
p.10 <- ggplot(data=data.set, aes(keystroke10, colour=name)) + 
  geom_histogram(fill="white")
p.10

p.20 <- ggplot(data=data.set, aes(keystroke20, colour=name)) + 
  geom_histogram(fill="white")
p.20

p.30 <- ggplot(data=data.set, aes(keystroke30, colour=name)) + 
  geom_histogram(fill="white")
p.30

p.40 <- ggplot(data=data.set, aes(keystroke40, colour=name)) + 
  geom_histogram(fill="white")
p.40

p.50 <- ggplot(data=data.set, aes(keystroke50, colour=name)) + 
  geom_histogram(fill="white")
p.50

p.60 <- ggplot(data=data.set, aes(keystroke60, colour=name)) + 
  geom_histogram(fill="white")
p.60

p.70 <- ggplot(data=data.set, aes(keystroke70, colour=name)) + 
  geom_histogram(fill="white")
p.70

p.80 <- ggplot(data=data.set, aes(keystroke80, colour=name)) + 
  geom_histogram(fill="white")
p.80

p.90 <- ggplot(data=data.set, aes(keystroke90, colour=name)) + 
  geom_histogram(fill="white")
p.90

p.100 <- ggplot(data=data.set, aes(keystroke100, colour=name)) + 
  geom_histogram(fill="white")
p.100
```

### Correlation matrix
```{r}
require(corrplot)
library(corrplot)
data.set$essay_typebin <- ifelse(data.set$essay_type == "True", 1, 0)
cor_data <- data.set[, 3:21]
corr_matrix <- cor(cor_data)
#simple correlation matrix
round(corr_matrix, 2)
#visualized
corrplot(corr_matrix, type = 'upper', order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

We will add keystokes over 110 to outliers later because there is a strong correlation
### Plot Data Construction
```{r}
val.vec <- c()
type.vec <- c()
sent.vec <- c()
theme.vec <- c()
for (i in 1:nrow(data.set)) {
  val <- max(data.set[i, 3:19])
  type <- data.set[i, 22]
  sent <- data.set[i, 21]
  theme <- data.set[i, 2]
  val.vec <- c(val.vec, val)
  type.vec <- c(type.vec, type)
  sent.vec <- c(sent.vec, sent)
  theme.vec <- c(theme.vec, theme)
}
plot.data <- data.frame(theme.vec, val.vec, type.vec, sent.vec)
colnames(plot.data) <- c('Theme', 'MaxKeystroke', "TypeofEssay", "Sentiment")

```

### Graphs on Dependencies between Maximum Keystroke Value and its sentiment, grouped by theme
```{r}
#Gay Marriage
ggplot(data=plot.data %>% filter(plot.data$Theme=='GayMerriage'), aes(MaxKeystroke, colour=TypeofEssay)) + geom_histogram() + ggtitle("Histogram of maximum keystroke over the essay on Gay Marriage")

ggplot(data=plot.data %>% filter(plot.data$Theme=='GayMerriage'), aes(x = MaxKeystroke, y =Sentiment, colour=TypeofEssay)) + geom_point() + ggtitle("Dependency of maximum keystroke over the essay on Gay Marriage and its sentiment")

#Gun Control
ggplot(data=plot.data %>% filter(plot.data$Theme=='GunControl'), aes(MaxKeystroke, colour=TypeofEssay)) + geom_histogram() + ggtitle("Histogram of maximum keystroke over the essay on Gun Control")

ggplot(data=plot.data %>% filter(plot.data$Theme=='GunControl'), aes(x = MaxKeystroke, y =Sentiment, colour=TypeofEssay)) + geom_point() + ggtitle("Dependency of maximum keystroke over the essay on Gun Control and its sentiment")

#ReviewAMT
ggplot(data=plot.data %>% filter(plot.data$Theme=='ReviewAMT'), aes(MaxKeystroke, colour=TypeofEssay)) + geom_histogram() + ggtitle("Dependency of maximum keystroke over the essay on Review and its sentiment")

ggplot(data=plot.data %>% filter(plot.data$Theme=='ReviewAMT'), aes(x = MaxKeystroke, y =Sentiment, colour=TypeofEssay)) + geom_point() + ggtitle("Histogram of maximum keystroke over the essay on Review")


```

People writing about Gay Marriage or Gun Control try to keep neutral position compared to essays on Review of the AMT where sentiment is either very negative or very positive.
Also we can say that huge keystroke almost never correspond to neutral opinion, so people take more time to write positive or negative thoughts.

### Simple Linear Model over Sentiment and Keystrokes
```{r}

model.1 <- lm(essay_sent ~ keystroke10 + keystroke20 + keystroke30 + keystroke30 + keystroke40 + 
                keystroke50 + keystroke60 + keystroke70 + keystroke80 + keystroke90 + keystroke100 + 
                keystroke110 + keystroke120 + keystroke130 + keystroke140 + keystroke150 + keystroke160
              + keystroke170 , data=data.set)
summary(model.1)

ggplot(data=data.set, aes(x = keystroke10 + keystroke20 + keystroke30 + keystroke30 + keystroke40 + 
                keystroke50 + keystroke60 + keystroke70 + keystroke80 + keystroke90 + keystroke100 + 
                keystroke110 + keystroke120 + keystroke130 + keystroke140 + keystroke150 + keystroke160
              + keystroke170,  y= essay_sent)) + geom_point() + geom_smooth(method = 'lm')
model_pred <- predict(model.1)
model.res <- data.set$essay_sent - model_pred
plot(model_pred, model.res)
abline(h=0)
```

Fighting the problem of too many variables, because they become insignificant. Let`s sum over keystroke

### Sum over keystrokes from 10 to 90
```{r}

vec <- c()
for (i in 1:nrow(data.set)) {
  vec <- c(vec, sum(data.set[i, 3:10]))
}
data.set$sum.10.80 <- vec 

```

### Filtering for treshold
```{r}
ind <- 90
for (i in 11:14) {
  ind.str <- paste(toString(ind),"bin",sep="")
  treshold <- mean(data.set[, i]) - 0.4*mean(data.set[, i])
  data.set[, i] <- ifelse(data.set[, i] >= treshold, 1, 0)
  ind <- ind + 10
}

### adding keystroke >= 120 to outliers
for ( i in 1:nrow(data.set)) {
  data.set[i, 20] <- data.set[i, 20] + sum(data.set[i, 14:19])
}

```
```{r}

model.2 <- lm(essay_sent ~ sum.10.80 + keystroke90 + keystroke100 + 
                keystroke110 + ouliers, data=data.set)
summary(model.2)

```

### Poisson Model
```{r}
library(rsq)
library(ggplot2)
poisson.model <- glm(essay_sent ~ keystroke10 + keystroke20 + keystroke30 + keystroke30 + keystroke40 + 
                keystroke50 + keystroke60 + keystroke70 + keystroke80 + keystroke90 + keystroke100 + 
                keystroke110 + keystroke120 + keystroke130 + keystroke140 + keystroke150 + keystroke160
              + keystroke170, data=data.set, family = poisson(link = "log"))
summary(poisson.model)
rsq(poisson.model,adj=TRUE)

poisson.model <- glm(essay_sent ~ sum.10.80 + keystroke90 + keystroke100 + 
                keystroke110 + ouliers, data=data.set, family = poisson(link = "log"))
summary(poisson.model)
rsq(poisson.model,adj=TRUE)

plot(predict(poisson.model), data.set$essay_sent, xlab="predicted",ylab="actual")
 abline(a=0,b=1)
```

### Logistic regression for sentiment
```{r}
data.set$essay_sentbin <- ifelse(data.set$essay_sent >= 5, 1, 0)
logistic.model <- glm(essay_sentbin ~ sum.10.80 + keystroke90 + keystroke100 + 
                keystroke110 + ouliers , data=data.set, family = "binomial")
summary(logistic.model)
rsq(logistic.model,adj=TRUE)

plot(model_pred, data.set$essay_sentbin)
model_pred <- predict(logistic.model)
model.res <- data.set$essay_sentbin - model_pred
plot(model_pred, model.res)
abline(h=0)
```

### Logistic regression for fake/true prediction
```{r}
logistic.model.type <- glm(essay_typebin ~ sum.10.80 + ouliers + factor(name) , data=data.set, family = "binomial")
summary(logistic.model.type)
rsq(logistic.model.type,adj=TRUE)
```

### Scaling the data
```{r}
data.set.scaled <- data.set %>% mutate_at(c("keystroke10", "keystroke20", "keystroke30", "keystroke40", "keystroke50", "keystroke60", "keystroke70", "keystroke80", "keystroke90", "keystroke100", "keystroke110", "keystroke120", "keystroke130", "keystroke140", "keystroke150", "keystroke160", "keystroke170"), ~(scale(.) %>% as.vector))
```

Doesn`t work for us, because we get negative values for keystrokes

### Ordinal Logistic Regression
```{r}
library(MASS)
summary(data.set)

data.set$essay_sent3 <- ifelse(data.set$essay_sent >= 6, "positive", ifelse(data.set$essay_sent <4, "negative", "neutural"))

model.polr <- polr(factor(essay_sent3)~essay_typebin + keystroke90 + keystroke100 + keystroke110 + keystroke120 + factor(name), data = data.set, Hess = TRUE)
summary(model.polr)
(ctable <- coef(summary(model.polr)))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
(ctable <- cbind(ctable, "p value" = p))
```

### Filtering the data
Here i deleted every observation where number of keystrokes was less than on 0.95 conf interval
```{r}
library(dplyr)

data.set.filtered <- data.set %>% filter(data.set$keystroke10 <= quantile(data.set$keystroke10, c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke20 <= quantile(data.set.filtered$keystroke20,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke30 <= quantile(data.set.filtered$keystroke30,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke40 <= quantile(data.set.filtered$keystroke40,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke50 <= quantile(data.set.filtered$keystroke50,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke60 <= quantile(data.set.filtered$keystroke60,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke70 <= quantile(data.set.filtered$keystroke70,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke80 <= quantile(data.set.filtered$keystroke80,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke90 <= quantile(data.set.filtered$keystroke90,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke100 <= quantile(data.set.filtered$keystroke100,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke110 <= quantile(data.set.filtered$keystroke110,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke120 <= quantile(data.set.filtered$keystroke120,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke130 <= quantile(data.set.filtered$keystroke130,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke140 <= quantile(data.set.filtered$keystroke140,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke150 <= quantile(data.set.filtered$keystroke150,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke160 <= quantile(data.set.filtered$keystroke160,c(.95)))
data.set.filtered <- data.set.filtered %>% filter(data.set.filtered$keystroke170 <= quantile(data.set.filtered$keystroke170,c(.95)))

```

```{r}
p.10 <- ggplot(data=data.set.filtered, aes(keystroke10, colour=name)) + 
  geom_histogram(fill="white")
p.10

p.20 <- ggplot(data=data.set.filtered, aes(keystroke20, colour=name)) + 
  geom_histogram(fill="white")
p.20

p.30 <- ggplot(data=data.set.filtered, aes(keystroke30, colour=name)) + 
  geom_histogram(fill="white")
p.30

p.40 <- ggplot(data=data.set.filtered, aes(keystroke40, colour=name)) + 
  geom_histogram(fill="white")
p.40

p.50 <- ggplot(data=data.set.filtered, aes(keystroke50, colour=name)) + 
  geom_histogram(fill="white")
p.50

p.60 <- ggplot(data=data.set.filtered, aes(keystroke60, colour=name)) + 
  geom_histogram(fill="white")
p.60

p.70 <- ggplot(data=data.set.filtered, aes(keystroke70, colour=name)) + 
  geom_histogram(fill="white")
p.70

p.80 <- ggplot(data=data.set.filtered, aes(keystroke80, colour=name)) + 
  geom_histogram(fill="white")
p.80

p.90 <- ggplot(data=data.set.filtered, aes(keystroke90, colour=name)) + 
  geom_histogram(fill="white")
p.90

```

### Simple Linear Model over Sentiment and Keystrokes
```{r}

model.1 <- lm(essay_sent ~ keystroke10 + keystroke20 + keystroke30 + keystroke30 + keystroke40 + 
                keystroke50 + keystroke60 + keystroke70 + keystroke80 + keystroke90 + keystroke100 + 
                keystroke110 + keystroke120 + keystroke130 + keystroke140 + keystroke150 + keystroke160
              + keystroke170 , data=data.set.filtered)
summary(model.1)

ggplot(data=data.set.filtered, aes(x = keystroke10 + keystroke20 + keystroke30 + keystroke30 + keystroke40 + 
                keystroke50 + keystroke60 + keystroke70 + keystroke80 + keystroke90 + keystroke100 + 
                keystroke110 + keystroke120 + keystroke130 + keystroke140 + keystroke150 + keystroke160
              + keystroke170,  y= essay_sent)) + geom_point() + geom_smooth(method = 'lm')
model_pred <- predict(model.1)
model.res <- data.set.filtered$essay_sent - model_pred
plot(model_pred, model.res)
abline(h=0)
```

Fighting the problem of too many variables, because they become insignificant. Let`s sum over keystroke

### SUm over keystrokes from 10 to 90
```{r}

vec <- c()
for (i in 1:nrow(data.set.filtered)) {
  vec <- c(vec, sum(data.set.filtered[i, 3:10]))
}
data.set.filtered$sum.10.80 <- vec 

```

### Filtering for treshold
```{r}
ind <- 90
for (i in 11:14) {
  ind.str <- paste(toString(ind),"bin",sep="")
  treshold <- mean(data.set.filtered[, i]) - 0.4*mean(data.set.filtered[, i])
  data.set.filtered[, i] <- ifelse(data.set.filtered[, i] >= treshold, 1, 0)
  ind <- ind + 10
}

### adding keystroke >= 120 to outliers
for ( i in 1:nrow(data.set.filtered)) {
  data.set.filtered[i, 20] <- data.set.filtered[i, 20] + sum(data.set.filtered[i, 14:19])
}

```

### Poisson Model
```{r}
library(rsq)
library(ggplot2)
poisson.model <- glm(essay_sent ~ sum.10.80 + keystroke90 + keystroke100 + 
                keystroke110 + ouliers + factor(name), data=data.set.filtered, family = poisson(link = "log"))
summary(poisson.model)
rsq(poisson.model,adj=TRUE)

plot(predict(poisson.model), data.set.filtered$essay_sent, xlab="predicted",ylab="actual")
 abline(a=0,b=1)
```

### Logistic regression for sentiment
```{r}
data.set.filtered$essay_sentbin <- ifelse(data.set.filtered$essay_sent >= 5, 1, 0)
logistic.model <- glm(essay_sentbin ~ sum.10.80 + keystroke90 + keystroke100 + 
                keystroke110 + ouliers , data=data.set.filtered, family = "binomial")
summary(logistic.model)

```

### Logistic regression for fake/true prediction
```{r}
logistic.model.type <- glm(essay_typebin ~ sum.10.80 + ouliers , data=data.set.filtered, family = "binomial")
summary(logistic.model.type)
rsq(logistic.model.type)
```

### Ordinal Logistic Regression
```{r}
library(MASS)
summary(data.set)

data.set.filtered$essay_sent3 <- ifelse(data.set.filtered$essay_sent >= 6, "positive", ifelse(data.set.filtered$essay_sent <4, "negative", "neutural"))

model.polr <- polr(factor(essay_sent3)~essay_typebin + keystroke90 + keystroke100 + keystroke110 + keystroke120 + factor(name), data = data.set.filtered, Hess = TRUE)
summary(model.polr)
(ctable <- coef(summary(model.polr)))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
(ctable <- cbind(ctable, "p value" = p))
```


### Random Forest for variables selection
```{r}
library(randomForest)
randForest.selection <- randomForest(factor(data.set.filtered$essay_sent3)~., data= data.set.filtered[, c(3:20)])
importance(randForest.selection)
```

